An efficient algorithm for integer lattice reduction

### 2023-06-30 tex/paper.tex
clarifying

To construct the initial basis vectors being reduced
as related to a finite field whose order is a large prime integer $q$,
the examples consider $q = 2^{13} - 1$ and $q = 2^{31} - 1$
(two well-known Mersenne primes).
The matrix whose columns are the initial basis vectors for the examples
takes the form
%
\begin{equation}
\label{matexample}
A = \left( \begin{array}{c|c} R & q \cdot \Id \\\hline
                              \Id & 0 \end{array} \right),
\end{equation}
%
where ``$\Id$'' denotes the identity matrix,
``$0$'' denotes the matrix whose entries are all zeros,
and $R$ denotes the random matrix whose entries are drawn
independent and identically distributed from the uniform distribution
over the integers $-(q-1)/2$, $-(q-3)/2$, \dots, $(q-3)/2$, $(q-1)/2$.
This special form of matrix effectively views the entries of $R$ modulo $q$,
as unimodular integer linear transformations acting on $A$ from the right
can add any integer multiple of $q$ to any entry of $R$.
The captions to the figures give the values of $q$ considered.

@@

The implementation of the LLL algorithm of~\cite{lenstra-lenstra-lovasz}
used here is entirely in IEEE standard double-precision arithmetic,
with acceleration via basic linear algebra subroutines (BLAS)
and enhanced accuracy via re-orthogonalization.
BLAS originates from~\cite{lawson-hanson-kincaid-krogh},
\cite{blas}, and others, with implementation in Apple's Accelerate Framework
of Xcode for MacOS Ventura 13.4 in the experiments reported here
(and compiled using Clang LLVM with the highest optimization flag,
{\tt -O3}, set). All experiments ran on a 2020 MacBook Pro
with a 2.3~GHz quad-core Intel Core i7 processor
and 3.733~GHz LPDDR4X SDRAM.
For the greatest portability, the implementation calls {\tt random()}
to generate random numbers; many compiler distributions
implement {\tt random()} via the BSD linear-congruential generator
(such compilers include our implementation's defaults
of MacOS Clang referenced by {\tt gcc} under Xcode
and GNU {\tt gcc} under Linux).

>>>>

To construct the initial basis vectors being reduced
as related to a finite field whose order is a large prime integer $q$,
the examples consider $q = 2^{13} - 1$ and $q = 2^{31} - 1$
(two well-known Mersenne primes).
The matrix whose columns are the initial basis vectors for the examples
takes the form
%
\begin{equation}
\label{matexample}
A = \left( \begin{array}{c|c} R & q \cdot \Id \\\hline
                              \Id & 0 \end{array} \right),
\end{equation}
%
where ``$\Id$'' denotes the identity matrix,
``$0$'' denotes the matrix whose entries are all zeros,
and $R$ denotes the pseudorandom matrix whose entries are drawn
independent and identically distributed from the uniform distribution
over the integers $-(q-1)/2$, $-(q-3)/2$, \dots, $(q-3)/2$, $(q-1)/2$.
This special form of matrix effectively views the entries of $R$ modulo $q$,
as unimodular integer linear transformations acting on $A$ from the right
can add any integer multiple of $q$ to any entry of $R$.
The captions to the figures give the values of $q$ considered.

@@

The implementation of the LLL algorithm of~\cite{lenstra-lenstra-lovasz}
used here is entirely in IEEE standard double-precision arithmetic,
with acceleration via basic linear algebra subroutines (BLAS)
and enhanced accuracy via re-orthogonalization.
BLAS originates from~\cite{lawson-hanson-kincaid-krogh},
\cite{blas}, and others, with implementation in Apple's Accelerate Framework
of Xcode for MacOS Ventura 13.4 in the experiments reported here
(and compiled using Clang LLVM with the highest optimization flag,
{\tt -O3}, set). All experiments ran on a 2020 MacBook Pro
with a 2.3~GHz quad-core Intel Core i7 processor
and 3.733~GHz LPDDR4X SDRAM.
For the greatest portability, the implementation calls {\tt random()}
to generate random numbers; many compiler distributions
implement {\tt random()} via the BSD linear-congruential generator
(such compilers include our implementation's defaults
of MacOS Clang referenced by {\tt gcc} under Xcode
and GNU {\tt gcc} under Linux). The results reported here pertain
to this class of pseudorandom numbers.
